---
title: "alex_reefcheck_fish"
format: html
---

Packages

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gtable)
library(grid)
library(scales)
library(tidyverse)
library(readr)
library(ggpattern)
```

initial data import

```{r}
fish_data <- read_csv("C:/R Projects/2025/port-report/data/Fish.csv")
kelp_data <- read_csv("C:/R Projects/2025/port-report/data/Kelp.csv")
invert_data <- read_csv("C:/R Projects/2025/port-report/data/Inverts.csv")
```

initial cleaning

```{r}
select_fish_data <- fish_data %>%
  mutate(year = year(mdy(date_start))) %>%
  select(site_name, year, transect, species, amount) %>%
  filter(site_name %in% c("Blake Island South", "Grain Terminal", "Freshwater Bay", "Magnolia", "Seattle Waterfront", "Wing Point")) %>%
  filter(year %in% c(2024, 2025))%>%
   mutate(site_name = case_when(
    site_name == "Grain Terminal" ~ "Centennial Park",
    TRUE ~ site_name  # Keep all other names as is
  ))

kelp_data <- kelp_data %>%
  mutate(year = year(mdy(date_start))) %>%
  select(site_name, year, transect, species, amount, distance) %>%
  filter(site_name %in% c("Blake Island South", "Grain Terminal", "Freshwater Bay", "Magnolia", "Seattle Waterfront", "Wing Point")) %>%
  filter(year %in% c(2024, 2025))

invert_data <- invert_data %>%
  mutate(year = year(mdy(date_start))) %>%
  select(site_name, year, transect, species, amount, distance) %>%
  filter(site_name %in% c("Blake Island South", "Grain Terminal", "Freshwater Bay", "Magnolia", "Seattle Waterfront", "Wing Point")) %>%
  filter(year %in% c(2024, 2025))
```

species richness formats
```{r}
fish_richness <- select_fish_data %>%
  group_by(site_name, species, year) %>%
  summarise(amount = sum(amount, na.rm = TRUE), .groups = "drop")
kelp_richness <- kelp_data %>%
  group_by(site_name, species, year) %>%
  summarise(amount = sum(amount, na.rm = TRUE), .groups = "drop")
invert_richness <- invert_data %>%
  group_by(site_name, species, year) %>%
  summarise(amount = sum(amount, na.rm = TRUE), .groups = "drop")
```

now moving on to fish only for graphs...

Figure 29 - Centennial Park only. Abundance (or density, TBD) comparing '24/'25, stacking shallow/deep perhaps, by species
```{r}
CP_fish_data <- select_fish_data %>%
  filter(site_name %in% c("Centennial Park")) %>%
  complete(year, species, site_name, transect, fill = list(amount = 0))
    


CP_fish_transect_data <- CP_fish_data %>%
  group_by(year, species, transect)%>%
  summarise(
    transect_amount = sum(amount, na.rm=TRUE),
    transect_density = transect_amount/120,
    .groups="drop")%>%
     mutate(
    depth_stratum = case_when(     
      transect %in% c(1:3, 7:12) ~ "Deep",
      transect %in% c(4:6, 13:18) ~ "Shallow",
      TRUE ~ NA_character_))

CP_fish_stratum_data <- CP_fish_transect_data %>%
  group_by(year, species, depth_stratum)%>%
  summarise(
    stratum_amount = sum(transect_amount, na.rm=TRUE),
    stratum_density = stratum_amount/9,
    stratum_cubic_density = stratum_density/120,
    transect_sd = sd(transect_amount),
    transect_density_sd = sd(transect_density),
    .groups="drop")

CP_fish_stratum_lowDensity <- CP_fish_stratum_data %>%
filter(!species %in% c("Kelp Perch", "Shiner Perch", "Striped Perch", "Pile Perch", "Forage Fish"))

CP_fish_stratum_lowDensity <- CP_fish_stratum_lowDensity %>%
  mutate(species = case_when(
    grepl("YOY", species) ~ "YOY Rockfish",  # Change any species with "YOY" in the name
    species == "Kelp Greenling (Male)" ~ "Kelp Greenling",
    TRUE ~ species  # Keep all other names as is
  ))


CP_fish_stratum_perchAndFGF <- CP_fish_stratum_data %>%
filter(species %in% c("Kelp Perch", "Shiner Perch", "Striped Perch", "Pile Perch", "Forage Fish"))

  CP_fish_stratum_perchAndFGF <- CP_fish_stratum_perchAndFGF %>%
  mutate(species = case_when(
    grepl("YOY", species) ~ "YOY Rockfish",  # Change any species with "YOY" in the name
    species == "Kelp Greenling (Male)" ~ "Kelp Greenling",
    TRUE ~ species  # Keep all other names as is
  ))


```

initial test plot - disregard
```{r}
ggplot(CP_fish_stratum_data, aes(x = species, y = stratum_amount, fill = interaction(depth_stratum, year))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("deep.2024" = "#1f77b4",      # dark blue
                               "shallow.2024" = "#aec7e8",    # light blue
                               "deep.2025" = "#ff7f0e",        # dark orange
                               "shallow.2025" = "#ffbb78"),    # light orange
                    labels = c("Deep 2024", "Shallow 2024", "Deep 2025", "Shallow 2025")) +
  labs(x = "Species",
       y = "Amount",
       fill = "Depth & Year",
       title = "Fish Abundance by Species, Depth, and Year at Grain Terminal") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate species names
        legend.position = "right")
```


low density fish plots (excluding perch & forage fish)
```{r}
RCF_CP_LD <- ggplot(CP_fish_stratum_lowDensity, 
       aes(y=species, x=stratum_density, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
    geom_vline(xintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9, reverse=TRUE),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  geom_errorbar(
    aes(xmin = pmax(stratum_density-transect_sd,0), xmax =stratum_density+transect_sd, 
                #group = interaction(depth_stratum, year)
        ),
    position = position_dodge(width=.9, reverse=TRUE),
    width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  xlab(bquote(Density (m^-2))) +
  theme_minimal() +
  theme(panel.grid.major.x = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.99,0.35),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),
        # Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.y = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.x = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_lowDensity$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(y="Species", x=expression(paste("Observations per 120m"^{3}," transect"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_x_continuous(expand=c(0,0, 0.05, 0),limits = c(-.05,2.35), breaks=seq(0,2.35,by=.5)) +
  scale_y_discrete(
    limits = rev(sort(unique(CP_fish_stratum_lowDensity$species))),  # Reverse alphabetical order
    labels = function(y) str_wrap(y, width = 11),
    expand = expansion(add = .66)) +
  coord_cartesian(xlim = c(-.05, 2.35))  #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_lowDensity.png",RCF_CP_LD, width=4,height=3, dpi = 500, scale = 1.75)
```
vertical version of above - yes this is the one to use.
```{r}
RCF_CP_LD_vert <- ggplot(CP_fish_stratum_lowDensity, 
       aes(x=species, y=stratum_density, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
    geom_hline(yintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  geom_errorbar(
    aes(ymin = pmax(stratum_density-transect_sd,0), ymax =stratum_density+transect_sd, 
                #group = interaction(depth_stratum, year)
        ),
    position = position_dodge(width=.9),
    width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  theme_minimal() +
  theme(panel.grid.major.y = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.99,0.45),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),
        # Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_lowDensity$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(x="Species", y=expression(paste("Observations per 120m"^{3}," transect"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_y_continuous(expand=c(0,0, 0.05, 0),limits = c(-.05,2.35), breaks=seq(0,2.35,by=.5)) +
  scale_x_discrete(
    #limits = rev(sort(unique(CP_fish_stratum_lowDensity$species))),  # Reverse alphabetical order
    labels = function(x) str_wrap(x, width = 11),
    expand = expansion(add = .66)) +
  coord_cartesian(ylim = c(-.05, 2.35))  #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_lowDensity.png",RCF_CP_LD_vert, width=4,height=3, dpi = 500, scale = 1.75)
```


high density fish plot (perch and forage fish)
```{r}
RCF_CP_HD <- ggplot(CP_fish_stratum_perchAndFGF, 
       aes(y=species, x=stratum_density, 
           fill=factor(year), 
           pattern = depth_stratum)) +
    geom_vline(xintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9, reverse=TRUE),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  geom_errorbar(
    aes(xmin = pmax(stratum_density-transect_sd,0), xmax =stratum_density+transect_sd, 
                #group = interaction(depth_stratum, year)
        ),
    position = position_dodge(width=.9, reverse=TRUE),
    width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  xlab(bquote(Density (m^-2))) +
  theme_minimal() +
   theme(panel.grid.major.x = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.99,0.35),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),# Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.y = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.x = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_perchAndFGF$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(y="Species", x=expression(paste("Observations per 120m"^{3}," transect"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_x_continuous(expand=c(0,0, 0.05, 0),limits = c(-1,45), breaks=seq(0,45,by=10)) +
  scale_y_discrete(
    limits = rev(sort(unique(CP_fish_stratum_perchAndFGF$species))),  # Reverse alphabetical order
    labels = function(y) str_wrap(y, width = 8),
    expand = expansion(add = .6)) +
  coord_cartesian(xlim = c(-1, 45))  #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_highDensity.png",RCF_CP_HD, width=4,height=3, dpi = 500, scale = 1.75)
```
vertical version of above.
```{r}
RCF_CP_HD_vert <- ggplot(CP_fish_stratum_perchAndFGF, 
       aes(x=species, y=stratum_density, 
           fill=factor(year), 
           pattern = depth_stratum)) +
    geom_hline(yintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  geom_errorbar(
    aes(ymin = pmax(stratum_density-transect_sd,0), ymax =stratum_density+transect_sd, 
                #group = interaction(depth_stratum, year)
        ),
    position = position_dodge(width=.9),
    width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  theme_minimal() +
   theme(panel.grid.major.y = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.99,0.48),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),# Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_perchAndFGF$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(x="Species", y=expression(paste("Observations per 120m"^{3}," transect"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_y_continuous(expand=c(0,0, 0.05, 0),limits = c(-1,45), breaks=seq(0,45,by=10)) +
  scale_x_discrete(
   # limits = rev(sort(unique(CP_fish_stratum_perchAndFGF$species))),  # Reverse alphabetical order
    labels = function(x) str_wrap(x, width = 8),
    expand = expansion(add = .6)) +
  coord_cartesian(ylim = c(-1, 45))  #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_highDensity.png",RCF_CP_HD_vert, width=4,height=3, dpi = 500, scale = 1.75)
```


low counts fish plot 
```{r}
RCF_CP_LC <- ggplot(CP_fish_stratum_lowDensity, 
       aes(y=species, x=stratum_amount, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
    geom_vline(xintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9, reverse=TRUE),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  #geom_errorbar(
  #  aes(xmin = pmax(stratum_density-transect_sd,0), xmax =stratum_density+transect_sd, 
                #group = interaction(depth_stratum, year)
  #      ),
  #  position = position_dodge(width=.9, reverse=TRUE),
  #  width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  theme_minimal() +
  theme(panel.grid.major.x = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.99,0.35),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),
        # Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.y = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.x = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_lowDensity$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(y="Species", x="Total Observations") +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_x_continuous(expand=c(0,0, 0.05, 0),limits = c(-.25,9.5), breaks=seq(0,9.5,by=2)) +
  scale_y_discrete(
    limits = rev(sort(unique(CP_fish_stratum_lowDensity$species))),  # Reverse alphabetical order
    labels = function(y) str_wrap(y, width = 11),
    expand = expansion(add = .66)) +
  coord_cartesian(xlim = c(-.25, 9.5))  #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_lowCounts.png",RCF_CP_LC, width=4,height=3, dpi = 500, scale = 1.75)
```
vertical version of above
```{r}
RCF_CP_LC_vert <- ggplot(CP_fish_stratum_lowDensity, 
       aes(x=species, y=stratum_amount, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
    geom_hline(yintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  #geom_errorbar(
  #  aes(xmin = pmax(stratum_density-transect_sd,0), xmax =stratum_density+transect_sd, 
                #group = interaction(depth_stratum, year)
  #      ),
  #  position = position_dodge(width=.9, reverse=TRUE),
  #  width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  theme_minimal() +
  theme(panel.grid.major.y = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  linewidth = 1),
        legend.position = c(0.99,0.45),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),
        # Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_lowDensity$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(x="Species", y="Total Observations") +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_y_continuous(expand=c(0,0, 0.05, 0),limits = c(-.25,9.5), breaks=seq(0,9.5,by=2)) +
  scale_x_discrete(
    #limits = rev(sort(unique(CP_fish_stratum_lowDensity$species))),  # Reverse alphabetical order
    labels = function(x) str_wrap(x, width = 11),
    expand = expansion(add = .66)) +
  coord_cartesian(ylim = c(-.25, 9.5))  #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_lowCounts.png",RCF_CP_LC_vert, width=4,height=3, dpi = 500, scale = 1.75)
```


high counts fish plot
```{r}
RCF_CP_HC <- ggplot(CP_fish_stratum_perchAndFGF, 
       aes(y=species, x=stratum_amount, 
           fill=factor(year), 
           pattern = depth_stratum)) +
    geom_vline(xintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9, reverse=TRUE),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  #geom_errorbar(
  #  aes(xmin = pmax(stratum_density-transect_sd,0), xmax =stratum_density+transect_sd, 
  #              #group = interaction(depth_stratum, year)
  #      ),
  #  position = position_dodge(width=.9, reverse=TRUE),
  #  width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  theme_minimal() +
   theme(panel.grid.major.x = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.99,0.35),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),# Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.y = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.x = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_perchAndFGF$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(y="Species", x="Total Observations") +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_x_continuous(expand=c(0,0, 0.05, 0),limits = c(-5,175), breaks=seq(0,175,by=40)) +
  scale_y_discrete(
    limits = rev(sort(unique(CP_fish_stratum_perchAndFGF$species))),  # Reverse alphabetical order
    labels = function(y) str_wrap(y, width = 8),
    expand = expansion(add = .6)) +
  coord_cartesian(xlim = c(-5, 190))  #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_highCounts.png",RCF_CP_HC, width=4,height=3, dpi = 500, scale = 1.75)
```
vertical version of above
```{r}
RCF_CP_HC_vert <- ggplot(CP_fish_stratum_perchAndFGF, 
       aes(y=species, x=stratum_amount, 
           fill=factor(year), 
           pattern = depth_stratum)) +
    geom_vline(xintercept = 0, 
             color = "gray", 
             linetype = "solid", 
             linewidth = 0.5) +
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  #geom_errorbar(
  #  aes(xmin = pmax(stratum_density-transect_sd,0), xmax =stratum_density+transect_sd, 
  #              #group = interaction(depth_stratum, year)
  #      ),
  #  position = position_dodge(width=.9, reverse=TRUE),
  #  width = 0.2) +
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  theme_minimal() +
   theme(panel.grid.major.y = element_line(color="gray", linewidth = 0.25),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.99,0.5),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),# Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.y = element_text(size = 12),         # Larger x-axis label
        axis.title.x = element_text(size = 12),         # Larger y-axis label
        axis.text.y = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.x = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  #geom_hline(yintercept = seq(0.5, length(unique(CP_fish_stratum_perchAndFGF$species)) - 0.5, by = 1), 
  #           color = "gray", 
  #           linetype = "solid", 
  #           linewidth = 0.5) +

  labs(y="Species", x="Total Observations") +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_x_continuous(expand=c(0,0, 0.05, 0),limits = c(-5,175), breaks=seq(0,175,by=40)) +
  scale_y_discrete(
    #limits = rev(sort(unique(CP_fish_stratum_perchAndFGF$species))),  # Reverse alphabetical order
    labels = function(y) str_wrap(y, width = 8),
    expand = expansion(add = .6)) +
  coord_cartesian(xlim = c(-5, 190))+
  coord_flip() #so error bars don't go below 0

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_CPonly_highCounts.png",RCF_CP_HC_vert, width=4,height=3, dpi = 500, scale = 1.75)
```


site comparisons: 

Step 1, data cleaning
```{r}
comp_fish_data <- select_fish_data %>%
  filter(transect <= 18, !species %in% c("No Fish Found"), !is.na(species))

comp_fish_data <- comp_fish_data %>%
  mutate(species = case_when(
    grepl("Rockfish", species) ~ "Rockfish",
    grepl("YOY", species) ~ "YOY Rockfish",  # Change any species with "YOY" in the name
    grepl("Perch", species) ~ "Perch",
   # is.na(species) ~ "No Fish Found",  # Convert NA to "No Fish Found"
    species %in% c("Buffalo Sculpin","Red Irish Lord", "Cabezon", "Whitespotted Greenling", "Kelp Greenling (Female)", "Lingcod", "Kelp Greenling (Male)", "Spiny Dogfish") ~ "Other Fish",
    #species == "No Fish Found" ~ "No Fish Found",  # Keep existing "No Fish Found"
    TRUE ~ species  # Keep all other names as is
  ))%>%
  complete(year, species, site_name, transect, fill = list(amount = 0))
    
comp_fish_transect_data <- comp_fish_data %>%
  group_by(year, site_name, species, transect)%>%
  summarise(
    transect_amount = sum(amount, na.rm=TRUE),
    transect_density = transect_amount/120,
    .groups="drop")%>%
     mutate(
    depth_stratum = case_when(     
      transect %in% c(1:3, 7:12) ~ "Deep",
      transect %in% c(4:6, 13:18) ~ "Shallow",
      TRUE ~ NA_character_),
    n_transects = case_when(
      site_name == "Seattle Waterfront" ~ 3,
      TRUE ~ 9
    ))

comp_fish_stratum_data <- comp_fish_transect_data %>%
  group_by(year, site_name, species, depth_stratum, n_transects)%>%
  summarise(
    stratum_amount = sum(transect_amount, na.rm=TRUE),
    transect_sd = sd(transect_amount),
    transect_density_sd = sd(transect_density),
    .groups="drop")

comp_fish_stratum_data <- comp_fish_stratum_data %>%
  mutate(
    stratum_density = stratum_amount/n_transects,
    stratum_cubic_density = stratum_density/120
  )

comp_fish_stratum_lowDensity <- comp_fish_stratum_data %>%
filter(!species %in% c("Perch", "Forage Fish"))

comp_fish_stratum_highDensity <- comp_fish_stratum_data %>%
filter(species %in% c("Perch", "Forage Fish"))

noFGF_comp_fish_stratum <- comp_fish_stratum_data %>%
  filter(!species %in% c("Forage Fish"))
```

Step 2: graphing

Density graphs first:

```{r}
species_order <- c("Rockfish", "YOY Rockfish", "Other Fish")
```

high density (FGF and Perch)
```{r}
ggplot(comp_fish_stratum_highDensity, 
       aes(x=species, y=stratum_density, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.025,
    pattern_angle = 45) +
  geom_errorbar(
    aes(ymin = pmax(stratum_density-transect_sd,0), ymax =stratum_density+transect_sd, 
       ),
    position = position_dodge(width=.9),
   width = 0.2) +  
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  facet_wrap(~ site_name, ncol = 3)+
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                    fill=NA,
                                    size = 1),
        axis.title.x = element_text(size = 12),         # Larger x-axis label
        axis.title.y = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 9),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  labs(x="Species", y = expression(paste("Observations per 120m"^{3}," transect"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
         pattern = guide_legend(title = "Depth Strata", order = 2,
                                override.aes = list(
                                  fill = "white",
                                  color = "black")))
  #coord_flip()

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_allSites_highDensity.png", width=4,height=2.5, dpi = 500, scale = 1.75)

```
low density (Rockfish, YOY, Other)
```{r}
ggplot(comp_fish_stratum_lowDensity, 
       aes(x=species, y=stratum_density, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.025,
    pattern_angle = 45) +
  geom_errorbar(
    aes(ymin = pmax(stratum_density-transect_sd,0), ymax =stratum_density+transect_sd, 
       ),
    position = position_dodge(width=.9),
   width = 0.2) +  
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
    scale_x_discrete(
    limits = species_order,  # Set the order
    labels = function(x) str_wrap(x, width = 9)  # Wrap after 10 characters
  ) +
  facet_wrap(~ site_name, ncol = 3)+
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                    fill=NA,
                                    size = 1),
        axis.title.x = element_text(size = 12),         # Larger x-axis label
        axis.title.y = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 9),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  labs(x="Species", y = expression(paste("Observations per 120m"^{3}," transect"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
         pattern = guide_legend(title = "Depth Strata", order = 2,
                                override.aes = list(
                                  fill = "white",
                                  color = "black")))
  #coord_flip()

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_allSites_lowDensity.png", width=4,height=2.5, dpi = 500, scale = 1.75)
```

now counts instead of density

high counts (Perch & FGF)
```{r}
ggplot(comp_fish_stratum_highDensity, 
       aes(x=species, y=stratum_amount, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.025,
    pattern_angle = 45) +
  #geom_errorbar(
  #  aes(ymin = pmax(stratum_density-transect_sd,0), ymax =stratum_density+transect_sd, 
  #     ),
  #  position = position_dodge(width=.9),
  # width = 0.2) +  
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  facet_wrap(~ site_name, ncol = 3)+
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                    fill=NA,
                                    size = 1),
        axis.title.x = element_text(size = 12),         # Larger x-axis label
        axis.title.y = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 9),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  labs(x="Species", y = expression(paste("Total Observations"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
         pattern = guide_legend(title = "Depth Strata", order = 2,
                                override.aes = list(
                                  fill = "white",
                                  color = "black")))+
  scale_y_continuous(expand=c(0,0, 0.05, 0),limits = c(-200,2750), breaks=seq(0,2750,by=500))

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_allSites_highCounts.png", width=4,height=2.5, dpi = 500, scale = 1.75)
```
and low counts (rockfish, YOY, other)
```{r}
ggplot(comp_fish_stratum_lowDensity, 
       aes(x=species, y=stratum_amount, 
           fill=factor(year), 
           pattern = depth_stratum)) + 
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = .9),
    color = "black",
    width = .9,
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.025,
    pattern_angle = 45) +
  #geom_errorbar(
  #  aes(ymin = pmax(stratum_density-transect_sd,0), ymax =stratum_density+transect_sd, 
  #     ),
  #  position = position_dodge(width=.9),
  # width = 0.2) +  
  scale_fill_manual(values=c("2024" = "#66B2FF", "2025" = "#B8E39B")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
      scale_x_discrete(
    limits = species_order,  # Set the order
    labels = function(x) str_wrap(x, width = 9)  # Wrap after 10 characters
  ) +
  facet_wrap(~ site_name, ncol = 3)+
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                    fill=NA,
                                    size = 1),
        axis.title.x = element_text(size = 12),         # Larger x-axis label
        axis.title.y = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 9),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  labs(x="Species", y = expression(paste("Total Observations"))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_Density = 0,
                               pattern_fill = NA,
                               pattern_Spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
         pattern = guide_legend(title = "Depth Strata", order = 2,
                                override.aes = list(
                                  fill = "white",
                                  color = "black")))#+
  #coord_flip()

ggsave("C:/R Projects/2025/port-report/figures/ReefCheckFish_allSites_lowCounts.png", width=4,height=2.5, dpi = 500, scale = 1.75)
```

extra stuff for now...
```{r}
#assiging depth strata to transects
working_fish_data <- select_fish_data %>%
  group_by(site_name, species, year, transect)%>%
  summarise(amount = sum(amount, na.rm=TRUE), .groups = "drop") %>%
  mutate(
    depth_stratum = case_when(     
      transect %in% c(1:3, 7:12) ~ "deep",
      transect %in% c(4:6, 13:18) ~ "shallow",
      TRUE ~ NA_character_),
    density = amount/120)


#group by depth stratum, year, species; sum amounts
fish_strata <- working_fish_data %>%
  group_by(site_name, species, year, depth_stratum) %>%
  summarise(
    amount=sum(amount, na.rm=TRUE),
    density = sum(density, na.rm=TRUE)/3,
    .groups="drop")

#check on completeness of data
fish_check <- select_fish_data %>%
  group_by(site_name, year) %>%
  mutate(transect_count = n_distinct(transect))
```

