---
title: "Kari's Script"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## *Here we gooooo....*

*\*\*This is a note for making the grey code section: ctrl + alt + i for this command!*

*Packages to install for this project:*

```{r}
library("tidyverse")
library("ggplot2")
library("readr")
library("dplyr")
library("scales")
library("lubridate")
library("stringr")
library("ggpattern")
library("googledrive")
library("googlesheets4")
library("patchwork")
```

*Pulling in the Reef Check data from 2024 and merging to 2025 + removing columns/rows that I don't want to keep - data clean-up*

```{r}
inverts_raw <- Inverts

#removing columns
inverts_clean <- inverts_raw %>% 
  select(site_name, date_start, date_end, transect, species, amount, distance)

#filtering rows that I want (site comparisons)
inverts_comparison_sites <- inverts_clean %>% 
  filter(site_name %in% c("Blake Island South", "Grain Terminal", "Freshwater Bay", "Magnolia", "Seattle Waterfront", "Wing Point"))

#updating the date column format and adding a year column
inverts_comparison_sites$date_start <- mdy(inverts_comparison_sites$date_start)
inverts_comparison_sites$date_end <- mdy(inverts_comparison_sites$date_end)
inverts_comparison_sites <- inverts_comparison_sites %>% 
  mutate(Year = year(date_start))

#mutating transect numbers to deep or shallow
inverts_comparison_sites <- inverts_comparison_sites %>% 
  mutate(depth_strata = ifelse(transect %in% c(1, 2, 3), "Deep",
                              ifelse(transect %in% c(4,5,6), "Shallow", NA)))

#changing Grain Terminal site name to Centennial Park in Site column
inverts_comparison_sites <- inverts_comparison_sites %>% 
  mutate(site_name = ifelse(site_name == "Grain Terminal", "Centennial Park", site_name))
```

*Comparing 2024 and 2025 inverts in the dataframe and completing the table with all transects being complete*

```{r}
#filtering out the years for comparison
yearly_comparison <- inverts_comparison_sites %>% 
  filter(site_name == "Centennial Park",
         Year %in% c(2024, 2025))

#adding in the missing zeros to the data, so each transect/species has three values to calculate the SD from

#creating the full table with all the missing pieces 
all_species <- unique(yearly_comparison$species)
all_years <- unique(yearly_comparison$Year)

transect_grid <- expand.grid(
  species = all_species,
  Year = all_years,
  depth_strata = c("Deep", "Shallow"),
  transect = 1:6
)

#filtering transect numbers by depth
transect_grid <- transect_grid %>% 
  mutate(
    transect = case_when(
      depth_strata == "Deep" & transect %in% 1:3 ~ transect,
      depth_strata == "Shallow" & transect %in% 4:6 ~ transect,
      TRUE ~ NA_real_)) %>% 
  filter(!is.na(transect))

#merging the data with the "mising data" <- fill in the blanks so all transects have all species, even if ZERO were observed for a transect
yearly_comparison_clean <- transect_grid %>% 
  left_join(yearly_comparison, by = c("species", "Year", "depth_strata", "transect")) %>% 
  mutate(amount = ifelse(is.na(amount), 0, amount))

```

*Calculating the density and SD (from original amount) for each invertebrate and adding to the same dataframe*

```{r}
#calculating SD and Density (per depth strata, hence the 180)
yearly_comparison_clean <- yearly_comparison_clean %>% 
  mutate(
    density = amount / 60)

compiled_yearly_comparison <- yearly_comparison_clean %>%   
  group_by(species, depth_strata, Year) %>% 
  summarise(
    total_amount = sum(amount),
    total_density = sum(density) / 3,
    SD = sd(density))
```

*Plotting inverts side-by-side for strata and including 2024 & 2025 data!*

```{r}
yearly_comparison <- ggplot(compiled_yearly_comparison, 
       aes(x=species, y=total_density, 
           fill=factor(Year), 
           pattern = depth_strata)) + 
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = 0.9), #, padding = 0.1, preserve = "single"),
    color = "black",
    width = 0.7,
    pattern_fill = "black",
    pattern_density = 0.2,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  geom_errorbar(
    aes(ymin = pmax(total_density-SD,0), ymax =total_density+SD), 
                #group = interaction(depth_strata, Year)),
    position = position_dodge(width=0.9), #, padding = 0.1, preserve = "single"),
    width = 0.2) +
  scale_fill_manual(values=c("2024" = "#EC7014", "2025" = "#FEC44F")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  ylab(bquote(Density (m^-2))) +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.98,0.02),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),
        # Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.x = element_text(size = 12),         # Larger x-axis label
        axis.title.y = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  labs(x="Species", y= expression(Density~(m^-2))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_density = 0,
                               pattern_fill = NA,
                               pattern_spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
  scale_y_continuous(expand=c(0,0),limits = c(0,0.8), breaks=seq(0,.8,by=0.2)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 14),
                   expand = expansion(add = 0.3)) +
  coord_cartesian(ylim = c(0, max(compiled_yearly_comparison$total_density + compiled_yearly_comparison$SD) * 1.1)) +  #so error bars don't go below 0
  coord_flip()

print(yearly_comparison)

#without mottled star
compiled_yearly_comparisonFILTERED <- compiled_yearly_comparison %>% 
  filter(!species %in% c("Mottled Star"))

#splitting species into two groups so the plot is easier to visualize
compiled_yearly_comparison$size_category <- ifelse(compiled_yearly_comparison$total_density > 0.15, "Large", "Small")

 ggplot(compiled_yearly_comparison, 
       aes(x=species, y=total_density, 
           fill=factor(Year), 
           pattern = depth_strata)) + 
  geom_bar_pattern(
    stat= "identity",
    position = position_dodge(width = 0.9), #padding = 0.15, preserve = "single"),
    color = "black",
    width = 0.85,
    pattern_fill = "black",
    pattern_density = 0.35,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  geom_errorbar(
    aes(ymin = pmax(total_density-SD,0), 
        ymax = total_density+SD),
        position = position_dodge(width=0.9),
        width = 0.2) +
  scale_fill_manual(values=c("2024" = "#EC7014", "2025" = "#FEC44F")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        legend.position = c(0.98,0.02),
        legend.justification = c(1,0),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),
        # Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(1, "lines"),
        axis.title.x = element_text(size = 12),         # Larger x-axis label
        axis.title.y = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10)) +       # Larger y-axis tick labels)
  labs(x="Species", y=expression(Density~(m^-2))) +
  guides(fill = guide_legend(title = "Year", order = 1,
                             override.aes = (list(
                               pattern = "none",
                               pattern_density = 0,
                               pattern_fill = NA,
                               pattern_spacing = 0,
                               pattern_angle = 0,
                               color = "black"))),
       pattern = guide_legend(title = "Depth Strata", order = 2,
                              override.aes = list(
                                fill = "white",
                                color = "black"
                              ))) +
   facet_wrap(~size_category, scales = "free_x", ncol = 1) +
   scale_y_continuous(expand=c(0,0),limits = c(0,0.8), breaks=seq(0,.8,by=0.1)) +
   coord_flip()
 

```

*Density of invertebrates, BUT with two panels to show high density species and low density species -\> using this for POS report for ease of viewing*

```{r}
#creating a panel variable
 species_separate <- c("Rock Crab", "Orange Cucumber", "Mottled Star", "Kelp Crab", "Kelp Crab (Juvenile)")
 
 compiled_yearly_comparison$panel <- ifelse(
   compiled_yearly_comparison$species %in% species_separate,
   "High Density Species",
   "Other Species")
 
 compiled_yearly_comparison$panel <- factor(
   compiled_yearly_comparison$panel,
   levels = c("High Density Species",
              "Other Species")
 )
 
 #breaking things into two plots and then putting back together
 compiled_yearly_comparison_selected <- subset(compiled_yearly_comparison,
                                               panel == "High Density Species")
 
 compiled_yearly_comparison_otherspecies <- subset(compiled_yearly_comparison, 
                                                   panel == "Other Species")

#create plot for high density species
high_density_species <- ggplot(
  compiled_yearly_comparison_selected, 
  aes(x = species, y = total_density, fill = factor(Year), pattern = depth_strata)) +
  geom_bar_pattern(
    stat = "identity",
    position = position_dodge(width = 0.9, reverse = TRUE), #, padding = 0, preserve = "single"),
    color = "black",
    width = 0.9,
    pattern_fill = "black",
    pattern_density = 0.35,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
  geom_errorbar(
    aes(ymin = pmax(total_density - SD, 0), ymax = total_density + SD),
    position = position_dodge(width=0.9, reverse = TRUE), #padding = 0.15, preserve = "single"),
    width = 0.2) +
  scale_fill_manual(values = c("2024" = "#66b2ff", "2025" = "#b8e39b")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  labs(x="Species", 
       y=expression(Density~(m^-2))) +
  scale_y_continuous(breaks = seq(0,1.4, by = 0.2),
                     expand = c(0,0)) +
  coord_flip(ylim = c(0, 1.4)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black", fill=NA, linewidth = 1),
        axis.text.x = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10),  
        axis.title.x = element_text(),
        axis.title.y = element_text(),
        legend.position = "none", #so both planels aren't competing for a legend
        strip.text = element_blank())
    
  #guides(
    #fill = guide_legend(
     # title = "Year",
      #override.aes = list(pattern = "none")),   # remove pattern from fill legend
    #pattern = guide_legend(
     # title = "Depth Strata",
      #override.aes = list(fill = "white")))     # make pattern legend background white

print(high_density_species)   
  

#create plot for all other species
all_other_species <- ggplot(
  compiled_yearly_comparison_otherspecies, 
  aes(x = species, y = total_density, fill = factor(Year), pattern = depth_strata)) + 
  geom_bar_pattern(
    stat = "identity",
    position = position_dodge(width = 0.9, reverse = TRUE), #padding = 0, preserve = "single"),
    color = "black",
    width = 0.9,
    pattern_fill = "black",
    pattern_density = 0.35,
    pattern_spacing = 0.02,
    pattern_angle = 45) +
 geom_errorbar(
    aes(ymin = pmax(total_density - SD, 0), ymax = total_density + SD),
    position = position_dodge(width=0.9, reverse = TRUE), #padding = 0.15, preserve = "single"),
    width = 0.2) +
  scale_fill_manual(values = c("2024" = "#66b2ff", "2025" = "#b8e39b")) +
  scale_pattern_manual(values = c("Deep" = "stripe", "Shallow" = "none")) +
  labs(x="Species", 
       y=expression(Density~(m^-2))) +
  scale_y_continuous(
    breaks = seq(0, max(compiled_yearly_comparison_otherspecies$total_density + compiled_yearly_comparison_otherspecies$SD), by = 0.02),
    expand = c(0,0)
  ) +
  coord_flip(
    ylim = c(0, max(compiled_yearly_comparison_otherspecies$total_density + compiled_yearly_comparison_otherspecies$SD))
  ) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(color="gray"),
        panel.grid.minor.x = element_blank(),
        panel.border = element_rect(color="black", fill=NA, linewidth = 1),
        axis.text.x = element_text(size = 10),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10),  
        axis.title.x = element_text(),
        axis.title.y = element_text(),
        legend.position = "right", 
        strip.text = element_blank()) +
  guides(
    fill = guide_legend(
      title = "Year",
      override.aes = list(pattern = "none")),   # remove pattern from fill legend
    pattern = guide_legend(
      title = "Depth Strata",
      override.aes = list(fill = "white")))     # make pattern legend background white

print(all_other_species)

(high_density_species / all_other_species) +
  plot_layout(axis_titles = "collect")

ggsave('C:/Users/kari/OneDrive/Data Work/Port of Seattle Report/port-report/figures/Densities of inverts present @ Centennial Park 2024-2025.jpg', 
       width = 6, height = 4, dpi = 400, scale = 1.75)

```

*Plotting the density of key kelp grazers at sites throughout the Salish Sea - comparison*

```{r}
inverts_comparison_sites_kelpgrazers <- inverts_comparison_sites %>%
  mutate(Year = as.numeric(Year),
         species = recode(species, 
                          "Green/Pallid Urchin" = "Green Urchin",
                          "Kelp Crab (Juvenile)" = "Kelp Crab (juv.)")) %>% 
  filter(species %in% c("Green Urchin", "Kelp Crab", "Kelp Crab (juv.)", 
                        "Purple Urchin", "Red Urchin"),
         Year %in% c(2024, 2025))

#completing data so all sites + transects have each species and a value (filling in zeros)
inverts_comparison_sites_kelpgrazers <- inverts_comparison_sites_kelpgrazers %>% 
  complete(
    site_name,
    species,
    Year,
    transect,
    fill = list(amount = 0))

#every site, year, transect has all species before calculations
inverts_comparison_sites_kelpgrazers <- inverts_comparison_sites_kelpgrazers %>%
  group_by(site_name, Year, transect) %>%
  complete(species, 
           fill = list(amount = 0, density_60m2 = 0, density_m2 = 0)) %>%
  ungroup()

#density + SD calculations
#density/m^2
inverts_comparison_sites_kelpgrazers <- inverts_comparison_sites_kelpgrazers %>% 
  mutate(density_m2 = amount / 60)


#mean density and SD by site, year, and species
inverts_comparison_sites_kelpgrazers <- inverts_comparison_sites_kelpgrazers %>%   
  group_by(site_name, species, Year) %>% 
  summarise(
    total_amount = sum(amount),
    total_density = sum(density_m2) / 6,
    SD = sd(density_m2))

#Plotted density of kelp grazers at sites throughout the Salish Sea
ggplot(inverts_comparison_sites_kelpgrazers, 
       aes(x = species, 
           y = total_density, 
           fill = factor(Year))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(
    ymin = pmax(total_density - SD, 0),
    ymax = total_density + SD),
    width = 0.2,
    position = position_dodge(width = 0.8)) +
  facet_wrap(~ site_name, scales = "free_x") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 5)) + 
  scale_fill_manual(values = c("2024" = "#66b2ff",
                               "2025" = "#b8e39b")) +
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,0.65), 
                     breaks = seq(0,0.65, by = 0.1)) +
  labs(
    x = "Species",
    y= expression(Density~(m^-2)),
    fill = "Year"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.major.y = element_line(color="gray"),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.border = element_rect(color="black",
                                    fill=NA,
                                    size = 1),
        legend.position = "right",
        legend.justification = "center",
        legend.box.just = "center",
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(4,4,4,4),
        # Increase legend text and key size
        legend.text = element_text(size = 10),          # Larger legend text
        legend.title = element_text(size = 11),         # Larger legend title
        legend.key.size = unit(1, "lines"),           # Larger legend keys
        legend.key.height = unit(0.8, "lines"),
        legend.key.width = unit(1, "lines"),
        axis.title.x = element_text(size = 12),         # Larger x-axis label
        axis.title.y = element_text(size = 12),         # Larger y-axis label
        axis.text.x = element_text(size = 8, hjust = 0.5),  # Larger x-axis tick labels
        axis.text.y = element_text(lineheight = 0.8, size = 10),
        strip.text = element_text(face = "bold"))


ggsave('C:/Users/kari/OneDrive/Data Work/Port of Seattle Report/port-report/figures/Invertebrate Density Site Comparisons- Kelp Predator Species.jpg', 
       width = 6, height = 4, dpi = 400, scale = 1.75)

```

Kelp Index data...cleaning and then filtering for each specific data that I want to use

```{r}
gs4_auth()
index_data_file <- drive_download(as_id("1ltFtDnOoMwLNBFVaQpp--U4_WzvAULDo"),
                                  path = tempfile(fileext = ".csv"))
index_data <- read_csv(index_data_file$local_path)

# clean up environment
rm(index_data_file)

#splitting datetime column into two columns and converting date to correct format
index_data <- index_data %>% 
  mutate(
    date = as.Date(datetime),
    time = format(datetime, "%H:%M:%S")
  )

#filtering by site
sites_to_keep <-
  c("centennialpark", "edmonds", "jeffersonhead", "wingpoint")

index_data_filtered <- index_data %>% 
  filter(site %in% sites_to_keep) 


#filter columns to keep
columns_to_keep <-
  c("site", 
    "position", 
    "pH", 
    "high_range_microsiemens_per_cm",
    "ph_temp_c",
    "con_temp_c",
    "date",
    "time",
    "abs_pres_kpa",
    "wl_logger_id",
    "con_logger_id",
    "wl_temp_c")

index_data_final <- index_data_filtered %>% 
  select(all_of(columns_to_keep))

#filtering out the data that doesn't overlap with Centennial Park
cutoff_Date <- as.Date("2024-09-19")
index_data_final <- index_data_final %>% 
  filter(date >= cutoff_Date)

#adding a year column ONLY
index_data_final <- index_data_final %>% 
  mutate(year = year(date))

#data is now ready with the sites that I want to use and the data (pH and conductivity)
```

pH

```{r}
#pH daily average surface and bottom by Site
daily_average_pH <-index_data_final %>% 
  group_by(site, date, position) %>% 
  summarise(
    mean_pH = mean(pH, na.rm = TRUE),
    n = n(),
    .groups = "drop")

daily_average_pH <- daily_average_pH %>% 
   mutate(year = year(date))

#removing rows with pH >9
daily_average_pH <- daily_average_pH %>%
  filter(mean_pH < 9)

#removing rows that have >96 and <96 readings
daily_average_pH <- daily_average_pH %>% 
  filter(n == 96)

# Filling in complete time series with NA where no data for plot
daily_average_pH <- daily_average_pH |> 
  group_by(site, position) |> 
  complete(date = seq(min(date), max(date), by = "1 day")) |>
  ungroup()

#making the x-axis of my plot into the annual quarters (Jan, Apl, Jul, Oct)
quarterly_breaks <- seq(
  from = floor_date(min(daily_average_pH$date), "year"),
  to = ceiling_date(max(daily_average_pH$date), "year"),
  by = "3 months")


#plotting two panel plot with surface and benthic position
#surface on top panel, benthic on bottom panel
daily_average_pH$position <- factor(
  daily_average_pH$position, 
  levels = c("surface", "bottom"))

ggplot(daily_average_pH,
       aes(x = date, y= mean_pH, color = site)) +
  geom_line(size = 1) +
  geom_point(size = 0.5) +
  facet_wrap(~position, 
             ncol = 1, 
             scales = "free_y",
             labeller = as_labeller((c(surface = "Surface", bottom = "Benthic")))) +
  scale_x_date(
    breaks = quarterly_breaks,
    date_labels="%b-%y") +
labs(x = "Month-Year",
     y = "Daily Average pH",
     color = "Site") +
  theme_minimal() +
  theme(panel.border=element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        panel.grid.major = element_line(color = "gray", size = 0.5),  # Major grid lines
        panel.grid.minor = element_line(color = "lightgray", size = 0.25),  # Minor grid lines
        axis.title.x = element_text(margin = margin(t = 10)),
        strip.text = element_text(face = "bold"),
        legend.position = c("right"),
        legend.justification = c("center"),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(5, 5, 5, 5), # Increase legend text and key size
        legend.text = element_text(size = 13),          # Larger legend text
        legend.title = element_text(size = 13),         # Larger legend title
        legend.key.size = unit(2, "lines"),           # Larger legend keys
        legend.key.height = unit(1.2, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(2, "lines"),
        text = element_text(size=14)) +
  scale_color_manual(values = c("edmonds" = "#6BAED6",
                                "wingpoint" = "#74C476",
                                "centennialpark"= "#FEC44F", 
                                "jeffersonhead" = "#C994C7"), # Custom colors 
                                labels = c("centennialpark" = "Centennial Park",
                                           "jeffersonhead" = "Jefferson Head",
                                           "edmonds" = "Edmonds",
                                           "wingpoint" = "Wing Point"))

#looking only at the weird Edmonds data
daily_average_pH_edmonds <- daily_average_pH %>%
  filter(site == "edmonds")


ggsave('C:/Users/kari/OneDrive/Data Work/Port of Seattle Report/port-report/figures/Daily Average pH @ Centennial Park with Site Comparisons 2024-2025.jpg', 
       width = 6, height = 4, dpi = 400, scale = 1.75)
                 
```

Conductivity –\> salinity

```{r}
#conductivity daily average surface and bottom
#converting microsiemens to salinity
install.packages("gsw") #R package for salinity conversion
library(gsw)
#citation of 'gsw' package
citation("gsw")


#converting microsiements to mS and kpac to dbar
index_data_final <- index_data_final %>% 
  mutate(
    pressure_dbar = if_else(position == "surface", 0, (abs_pres_kpa / 10) - 10.1325), #using seawater pressure constant 
    conductivity_mS = high_range_microsiemens_per_cm / 1000,
    salinity = case_when(site == "centennialpark" & 
                           con_logger_id == "CON1" & # CON1 temp readings bad - use WL temp readings
                           position == "bottom" ~
                           gsw_SP_from_C(conductivity_mS, wl_temp_c, pressure_dbar),
                         # Use temp from CON sensor for all others 
                         TRUE ~ gsw_SP_from_C(conductivity_mS, con_temp_c, pressure_dbar)))

#daily average salinity
daily_average_salinity <-index_data_final %>% 
  group_by(site, date, position) %>% 
  summarise(
    mean_salinity = mean(salinity, na.rm = TRUE),
    n = n(),
    .groups = "drop")

# Add in position = bottom for Edmonds NA positions
# Thinking that bad data from 9/2024 CON6 is the funky one here
daily_average_salinity <- daily_average_salinity |>
  mutate(position = if_else(is.na(position), "bottom", position))

daily_average_salinity <- daily_average_salinity %>% 
  mutate(year = year(date))

#removing rows that have >96 and <96 readings
daily_average_salinity <- daily_average_salinity %>% 
  filter(n == 96)

# Filling in complete time series with NA where no data for plot
daily_average_salinity <- daily_average_salinity |> 
  group_by(site, position) |> 
  complete(date = seq(min(date), max(date), by = "1 day")) |>
  ungroup()


#making the x-axis of my plot into the annual quarters (Jan, Apl, Jul, Oct)
quarterly_breaks <- seq(
  from = floor_date(min(daily_average_salinity$date), "year"),
  to = ceiling_date(max(daily_average_salinity$date), "year"),
  by = "3 months")


#plotting two panel plot with surface and bottom position
#surface on top panel, bottom on bottom panel
daily_average_salinity$position <- factor(
  daily_average_salinity$position, 
  levels = c("surface", "bottom"))

# Replace dates with "bad data" with NA
daily_average_salinity <- daily_average_salinity |>
  mutate(mean_salinity = 
           case_when(site == "edmonds" &
                       position == "surface" &
                       date >= as.Date("2024-10-24") &
                       date <= as.Date("2024-12-12") ~ NA_real_,
                     
                     site == "edmonds" &
                       position == "surface" &
                       date >= as.Date("2025-04-23") &
                       date <= as.Date("2025-07-13") ~ NA_real_, 
                     
                     site == "jeffersonhead" &
                       position == "surface" &
                       date >= as.Date("2025-04-27") &
                       date <= as.Date("2025-07-13") ~ NA_real_, 
                     
                     site == "wingpoint" &
                       position == "surface" &
                       date >= as.Date("2024-11-07") &
                       date <= as.Date("2025-03-13") ~ NA_real_, 
                     
                     site == "centennialpark" &
                       position == "bottom" &
                       date >= as.Date("2025-05-14") &
                       date <= as.Date("2025-06-11") ~ NA_real_, 
                     
                     site == "edmonds" &
                       position == "bottom" &
                       date >= as.Date("2025-05-04") &
                       date <= as.Date("2025-07-13") ~ NA_real_, 
                     
                     site == "jeffersonhead" &
                       position == "bottom" &
                       date >= as.Date("2025-06-11") &
                       date <= as.Date("2025-08-27") ~ NA_real_, 
                     
                     site == "jeffersonhead" &
                       position == "bottom" &
                       date >= as.Date("2025-09-29") ~ NA_real_, 
                     
                     site == "wingpoint" &
                       position == "bottom" &
                       date >= as.Date("2025-07-16") &
                       date <= as.Date("2025-07-31") ~ NA_real_, 
                     
                     site == "edmonds" &
                       position == "bottom" &
                       date >= as.Date("2025-09-1") ~ NA_real_, 
                     
                     TRUE ~ mean_salinity)
         )



ggplot(daily_average_salinity, 
       aes(x = date, y= mean_salinity, color = site)) +
  geom_line(size = 1) +
  geom_point(size = 0.5) +
  facet_wrap(~position, ncol = 1, scales = "free_y",
             labeller = as_labeller((c(surface = "Surface", bottom = "Benthic")))) +
  scale_x_date(
    breaks = quarterly_breaks,
    date_labels="%b-%y") +
labs(x = "Month-Year",
     y = "Daily Average Salinity (ppt)",
     color = "Site") +
  theme_minimal() +
  theme(panel.border=element_rect(color="black",
                                  fill=NA,
                                  size = 1),
        panel.grid.major = element_line(color = "gray", size = 0.5),  # Major grid lines
        panel.grid.minor = element_line(color = "lightgray", size = 0.25),  # Minor grid lines
        axis.title.x = element_text(margin = margin(t = 10)),
        strip.text = element_text(face = "bold"),
        legend.position = c("right"),
        legend.justification = c("center"),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin = margin(5, 5, 5, 5), # Increase legend text and key size
        legend.text = element_text(size = 13),          # Larger legend text
        legend.title = element_text(size = 13),         # Larger legend title
        legend.key.size = unit(2, "lines"),           # Larger legend keys
        legend.key.height = unit(1.2, "lines"),         # Adjust the height of legend keys (if necessary)
        legend.key.width = unit(2, "lines"),
        text = element_text(size=14)) +
  scale_color_manual(values = c("edmonds" = "#6BAED6",
                                "wingpoint" = "#74C476",
                                "centennialpark"= "#FEC44F", 
                                "jeffersonhead" = "#C994C7"), # Custom colors 
                                labels = c("centennialpark" = "Centennial Park",
                                           "jeffersonhead" = "Jefferson Head",
                                           "edmonds" = "Edmonds",
                                           "wingpoint" = "Wing Point"))

ggsave('C:/Users/kari/OneDrive/Data Work/Port of Seattle Report/port-report/figures/Daily Average salinity @ Centennial Park with Site Comparisons 2024-2025.jpg', 
       width = 6, height = 4, dpi = 400, scale = 1.75)

```







